Ms. No. PSRM-OA-2019-0126 entitled ``Taking Dyads Seriously''

### 2021-01-04 revisionMemo/rev2.tex

\begin{enumerate}
	\item The additive and multiplicative effects (AME) model presented in the paper allows for a better treatment of first, second and third order dependencies in the data, and account for unobserved sources of bias in the data. The presentation of the AME estimator is clear, and the simulations provide clear evidence that the model outperforms naive OLS/GLM models fitted to the same data. The paper will be of great interest to the PSRM audience and is likely to have a strong impact on future research in international relations. I do, however, have a few suggestions for revisions that will hopefully strengthen the paper’s contributions.
	\begin{itemize}
		\item \textcolor{blue}{ \emph{
		We thank the reviewer for their comments and suggestions.
		}}
	\end{itemize}
	\item The paper shows that the AME model performs well in comparison to the “oracle” model where all sources of dependencies in the data are accounted for, which is reassuring. Yet this comparison feels incomplete. First, the naive model is clearly misspecified given the underlying data generating process. A large number of dependencies in international relations are linked to geography. What would happen if the dependencies were modeled using spatial econometric techniques?
	\begin{itemize}
		\item \textcolor{blue}{ \emph{
			The reviewer raises an important connection to the spatial literature. The types of geographic dependencies in international relations are generally monadic in nature -- civil war spreads from Rwanda to Uganda, or protest from Lebanon to Syria. With dyadic data, however, we cannot use the common geographic adjacency matrices used for normal spatial econometrics, because in dyadic data a W matrix would need to specify how dyad i,j relates to dyad k,l (for every i,j,k,l). In the simulation, we do not include a spatial lag model because this would require specifying a spatial dependence pattern (see Plumper and Neumayer 2010 for a more in-depth discuss on spatial lags). \textbf{CD}: I added this as a footnote on p.9. I moved this footnote around a LOT but I think it needs to live somewhere in the simulation. It felt awkward within the text.
			}}
%		CDMG: Flag for discussion. We brainstormed how to fit this into the simulation, but wouldn't simulating a W matrix be a bit arbitrary here?
%		Reference papers that have thought about what happens when we get the wrong W
%		general gist .. spatial models of dyadic dvs are great if we know how to characterize the diffusion amtrices, but the point of our approach is that we may not know how to actually characterize the ways in which actors are related to one another on some diffusion space.
%		reference franzesse et al mstar framework as a way to operationalize multiple pathways of diffusion in a spatial econometric approach
	\end{itemize}
	\item It would help if the author discussed in more depth the properties of the AME estimator in the presence of unobserved dependencies. One would imagine that the oracle model, where all the unobserved factors and dependencies are accounted for should always outperform a parametric or semi-parametric estimator. Relatedly, it also would help to have a more technical discussion of the properties of AME models, particularly in terms of the bias-variance trade off.
	\begin{itemize}
		\item \textcolor{blue}{ \emph{
		add example of case where omitted variable is correlated
		tech discussion bias-variance

		\begin{figure}[ht]
		\caption{Regression parameter estimates for the standard, AME, and oracle models from 1,000 simulations. Summary statistics are presented through a traditional box plot, and the estimates from each simulation are visualized as well as points.}
		\label{fig:ameBias}
		\includegraphics[width=1\textwidth]{ameSimBias_all_corrProbitMed.pdf} \\
		\end{figure}

		\begin{figure}[ht]
		\caption{Regression parameter estimates for the standard, AME, and oracle models from 1,000 simulations. Summary statistics are presented through a traditional box plot, and the estimates from each simulation are visualized as well as points.}
		\label{fig:ameBias}
		\includegraphics[width=1\textwidth]{ameSimBias_all_corrProbitHi.pdf} \\
		\end{figure}

		\begin{figure}[ht]
		\caption{Regression parameter estimates for the standard, AME, and oracle models from 1,000 simulations. Summary statistics are presented through a traditional box plot, and the estimates from each simulation are visualized as well as points.}
		\label{fig:ameBias}
		\includegraphics[width=1\textwidth]{ameSimBias_all_corrMed.pdf} \\
		\end{figure}

		\begin{figure}[ht]
		\caption{Regression parameter estimates for the standard, AME, and oracle models from 1,000 simulations. Summary statistics are presented through a traditional box plot, and the estimates from each simulation are visualized as well as points.}
		\label{fig:ameBias}
		\includegraphics[width=1\textwidth]{ameSimBias_all_corrHi.pdf} \\
		\end{figure}

@@

		We've added additional discussion about how each of the variables in our simulation were generated. To make the model more applicable to the types of networks we study in IR we have provided a probit example.

@@

		The reviewer raises a valid point. We have expanded our tutorial in Appendix C, particularly in giving more detail and concrete examples when it comes to formatting the data, which may represent the biggest hurdle for some users. We also...

>>>>

\begin{enumerate}
	\item The additive and multiplicative effects (AME) model presented in the paper allows for a better treatment of first, second and third order dependencies in the data, and account for unobserved sources of bias in the data. The presentation of the AME estimator is clear, and the simulations provide clear evidence that the model outperforms naive OLS/GLM models fitted to the same data. The paper will be of great interest to the PSRM audience and is likely to have a strong impact on future research in international relations. I do, however, have a few suggestions for revisions that will hopefully strengthen the paper’s contributions.
	\begin{itemize}
		\item \textcolor{blue}{ \emph{
		We thank the reviewer for their comments and suggestions.
		}}
	\end{itemize}
	\item The paper shows that the AME model performs well in comparison to the “oracle” model where all sources of dependencies in the data are accounted for, which is reassuring. Yet this comparison feels incomplete. First, the naive model is clearly misspecified given the underlying data generating process. A large number of dependencies in international relations are linked to geography. What would happen if the dependencies were modeled using spatial econometric techniques?
	\begin{itemize}
		\item \textcolor{blue}{ \emph{
			The reviewer raises an important connection to the spatial literature. The types of geographic dependencies in international relations are generally monadic in nature -- civil war spreads from Rwanda to Uganda, or protest from Lebanon to Syria. With dyadic data, however, we cannot use the common geographic adjacency matrices used for normal spatial econometrics, because in dyadic data a W matrix would need to specify how dyad i,j relates to dyad k,l (for every i,j,k,l). In the simulation, we do not include a spatial lag model because this would require specifying a spatial dependence pattern (Plumper and Neumayer (2010) provide an in-depth discussion on varying operationalizations of spatial lags). If there is a particular type of spatial dependence that the reviewer would like us to simulate, however, we are happy to do so.
			}}
	\end{itemize}
	\item It would help if the author discussed in more depth the properties of the AME estimator in the presence of unobserved dependencies. One would imagine that the oracle model, where all the unobserved factors and dependencies are accounted for should always outperform a parametric or semi-parametric estimator. Relatedly, it also would help to have a more technical discussion of the properties of AME models, particularly in terms of the bias-variance trade off.
	\begin{itemize}
		\item \textcolor{blue}{ \emph{
			Estimation certainly gets more challenging when the omitted variable is correlated with the observed variable. In the case of normal linear regression, we can show that in general a regression coefficient is unbiased for its ``unconditional effect'', which is the direct effect plus something related to the direct effects of any omitted variables that it is correlated with. For example, suppose in our simulation setup $X_{i,j}$ is correlated with $W_{i,j}$ via the relation $W_{i,j} = \alpha X_{i,j} + e_{i,j}$. Then: }}

			\begin{align*}
			  Y_{i,j} & = \mu + \beta X_{i,j} + \gamma W_{i,j} + \epsilon_{i,j} \\
			  &= \mu + (\gamma \alpha+\beta ) X_{i,j} + (\gamma Z_{i,j} + \epsilon_{i,j})  \\
			  &= \mu + \tilde \beta X_{i,j} + \tilde \epsilon_{i,j}
			\end{align*}

		\item \textcolor{blue}{ \emph{
			The naive model is in some sense still ``correct'' as it is an unbiased estimator of the unconditional (direct plus indirect) effect $\gamma\alpha + \beta$ of $X$ on $Y$. However, there is no way to eliminate this extra bias without making further assumptions about the nature of the omitted variable. Additonally, in the binary case with the probit setup, other sources of bias get introduced as well because of the nonlinearity of the population moments as a function of the parameters. An interesting question is whether or not we can reduce the bias in the network setting by assuming the omitted variable $w_{i,j}$ is the product of latent node-specific factors. An idea related to this is discussed in \citet{minhas:etal:2017:arxiv}.}}

		\item \textcolor{blue}{ \emph{
			For now though to show the amount of bias that gets introduced into our setup when $X$ and $W$ are correlated we repeat our simulation exercise but induce varying levels of correlation between $X$ and $W$. Results with $X$ and $W$ correlated at 0.4 and then 0.7 are shown below, respectively.}}

			\begin{figure}[ht]
			\caption{Regression parameter estimates for the standard, AME, and oracle models from 1,000 simulations. Summary statistics are presented through a traditional box plot, and the estimates from each simulation are visualized as well as points. $X$ and $W$ are correlated at 0.4 in this case.}
			\label{fig:ameBias_corrMed}
			\includegraphics[width=1\textwidth]{ameSimBias_all_corrProbitMed.pdf} \\
			\end{figure}

			\begin{figure}[ht]
			\caption{Regression parameter estimates for the standard, AME, and oracle models from 1,000 simulations. Summary statistics are presented through a traditional box plot, and the estimates from each simulation are visualized as well as points. $X$ and $W$ are correlated at 0.7 in this case.}
			\label{fig:ameBias_corrHi}
			\includegraphics[width=1\textwidth]{ameSimBias_all_corrProbitHi.pdf} \\
			\end{figure}
			\FloatBarrier
			In terms of the bias-variance tradeoff discussion, this enters most prominently as a consideration for AME when choosing the dimensions of the latent factor space, specifically $K$ in the model. Everytime we add another dimension to the latent factor space the number of parameters in the model increases notably. For example, in a directed network with a choice of $K=1$ there are $2 \times n$ parameters introduced from the LFM portion of the model. With $K=2$ this number increases by another $2 \times n$. Increasing K certainly comes with a cost in terms of the number of extra parameters that are introduced into the model.

@@

		We thank the reviewer for this comment and have revised our simulation setup as a result. Most IR studies into networks have been interested in modeling binary outcomes (e.g., indicators for international agreements between countries or conflicts between countries), thus we we have revised our simulation to have a binary dependent variable. Key results remain the same.
		Additionally, we have thoroughly revised our description to make it clearer how each of the variables in the simulations were generated.

@@

		The reviewer raises a valid point. We have made a number of revisions and extensions to the tutorial at the end of the Appendix so that applied scholars can better take advantage of AME.
